---
title: "Data pipelines"
output: html_notebook
---

# Setup

In order to load the data, the *here* library is used to access the path with respect to the working directory of the project (CRIS_single-sample folder, which contains the .Rproj file). 

```{r setup}
library(here)
source(here('src','load_libraries.r'))
source(here('src','utils','source_utils.r'))
source(here('src','loader_writer','load_data.r'))
source(here('src','data_management','source_data_management.r'))
source(here('src','classifiers','source_classifiers.r'))
source(here('src','pipelines','source_pipelines.r'))
# Data loading
load_gmql_grch38(filtered = TRUE, uniformed = TRUE)

# Reference loading
NTP_ref <- load_file(path_loader$get_path("NTP_REF_TCGA"))
```

# CRISData class

In order to replicate the NTP classification, a *CRISData* object without no reference is required.
To load this type of data, it is sufficient to:

```{r}
cris_data_no_ref <- cris_no_ref_data_pipeline(GMQL_GRCH38_FILTERED$cpm, log2_tr = FALSE)
```

In order to apply TSP, a CRISData object with not-null reference is required. To load this type of data, it is sufficient to:

```{r}
cris_data <- cris_data_pipeline(expset = GMQL_GRCH38_FILTERED$cpm, 
                                ref    = NTP_ref$result,
                                conf   = 'conf', 
                                log2_tr = FALSE)
```

The *expset* parameter requires the data ExpressionSet; the *ref* parameter requires the output result of an NTP classification. The *conf* parameter can be either *conf*, *not_conf*, or *all*, to load respectively confident, not confident or all samples. A sample is considered confident if, for the assigned class, *BH.FDR < 0.2*. Finally, *log2_tr* parameter specifies if the expression must be transformed as *log2(1 + expression)*. 
For NTP and TSP, *log2tr* has been set to FALSE.

# SLData class

The SLData class is used to store data on which single-label classification will be computed.

```{r}
sldata <- sl_data_pipeline(expset = GMQL_GRCH38_FILTERED$cpm,
                           ref    = NTP_ref$result,
                           conf   = 'conf',
                           train_samples = c(),
                           log2_tr = TRUE)
```

The *expset*,*ref*,*conf*,*log2_tr* parameters are analogue to the ones specified for *CRISData* object (actually, *sldata* object uses a *CRISData* within its instantiation). The *train_samples* parameter is used to specify whether the dataset must be split or not.
When *train_samples* parameter is equal to *c()*, all samples are put into testing. When *train_samples* is a not-empty character vector (vector of aliquot IDs), the specified samples are put into training dataset:

```{r}
sldata <- sl_data_pipeline(expset = GMQL_GRCH38_FILTERED$cpm,
                           ref    = NTP_ref$result,
                           conf   = 'conf',
                           train_samples = load_training_samples(),
                           log2_tr = TRUE)
```

In order to access train and test data, it is sufficient to:

```{r}
head(sldata$train_)
head(sldata$test_)
```

In order to access the reference (predicted class, NTP correlation scores and FDR for all classes), it is sufficient to:

```{r}
head(sldata$train_ref)
head(sldata$test_ref)
```

Each sample is referenced through its aliquot-id. The predicted class is represented in *predicted.label2* attribute; the NTP correlation scores (computed as 1 - distances) are represented in *CRIS-A, CRIS-B, ... CRIS-E* attributes; the BH.FDR values are represented in *BH.FDR.CRIS.A, ... , BH.FDR.CRIS.E*. 
*perc* and *seed* attributes are inherited from generic class *Data*, but not used. The entire dataset and the entire reference can be examined through:

```{r}
head(sldata$data)
head(sldata$ref)
```

# MLData class

The MLData class is used to store data on which multi-label classification will be computed. The parameters are the same of *sl_data_pipeline*; in addition, however, the *dist_thr* threshold is necessary to transform the single-label NTP reference into a multi-label NTP reference. This parameter (whose name stands for *distance thresholds*) is a vector of 5 doubles (one threshold for each CRIS class).
Thresholds can be accessed with:

```{r}
ntp_thr <- load_file(path_loader$get_path('NTP_THR'))
```

Thresholds are specific to each dataset (TCGA or PDX). In order to get the thresholds for TCGA, it is sufficient to:

```{r}
ntp_thr$tcga
```

Similarly, for PDX:

```{r}
ntp_thr$pdx
```

Now that thresholds have been loaded, MLData object can be created:

```{r}
mldata <- ml_data_pipeline(expset = GMQL_GRCH38_FILTERED$cpm,
                           ref    = NTP_ref$result,
                           conf   = 'conf',
                           dist_thr = ntp_thr$tcga,
                           train_samples = c(),
                           log2_tr = TRUE)
```

As before, to split the data between training and testing:

```{r}
mldata <- ml_data_pipeline(expset = GMQL_GRCH38_FILTERED$cpm,
                           ref    = NTP_ref$result,
                           conf   = 'conf',
                           dist_thr = ntp_thr$tcga,
                           train_samples = load_training_samples(),
                           log2_tr = TRUE)
```

The entire data (which is an *mldr* object) and the entire reference can be accessed through

```{r}
head(mldata$data$dataset)
head(mldata$ref)
```

# PREPARED TCGA AND PDX DATA

In order to load confident TCGA data splitted into training and testing for single-label classifiers it is sufficient to (with CRIS signature as feature set):

```{r}
sldata <- load_prepared_tcga_data(confident = 'conf',
                                  uniformed = TRUE,
                                  fs_type = 'ntp_only',
                                  type = 'sl',
                                  samples_filter = NULL,
                                  load_training = TRUE)
```

The *confident* parameter, which be either *conf*, *not_conf*, *all* selects respectively confident, not confident or all data. confident samples are characterized by BH.FDR of the assigned class below 0.02.
The *uniformed* parameter (either *true* or *false*), tells if the NTP genes must be aligned to the reference used in PDX\_GRCh38 or not (we advise to keep it always to *true*). The *fs_type* parameter allows to specify the feature set. In order to use the original CRIS signature, it should be set to *ntp_only*. If set to *c("bio_driven","lasso")* all the genes surviving to the biological filter and to the lasso step can be obtained. The *type* parameter specifies the type of data: *cris_no_ref* prepares CRISData for NTP replication; *cris* prepares CRISData for TSP replication; *sl* prepares data for single-label classification. *ml* prepares data for multi-label classifier.
The *samples_filter* parameter can contain vector of aliquot ID to be filtered out by the available data, before possibly splitting in training and testing. By default, it is set to *NULL*.
The *load_training* parameter can be set to *TRUE* or to *FALSE*. If *TRUE*, data are splitted into training and testing, otherwise they are all put into testing.

For example, to obtain all the confident TCGA data splitted into training and testing for multi-label problem transformation classifiers (with CRIS signature as feature set):


```{r}
mldata <- load_prepared_tcga_data(confident = 'conf',
                                  uniformed = TRUE,
                                  fs_type = 'ntp_only',
                                  type = 'ml',
                                  samples_filter = NULL,
                                  load_training = TRUE)
```

In order to load testing TCGA data for TSP it is sufficient to:

```{r}
tsp_data <- load_prepared_tcga_data(confident = 'conf',
                                  uniformed = TRUE,
                                  fs_type = 'ntp_only',
                                  type = 'cris',
                                  samples_filter = load_tcga_testing_samples(),
                                  load_training = FALSE)
```

NB: in order to apply TSP, the only usable feature set is *ntp_only*. In order to load PDX confident data for single-label classifiers:
For PDX, the approach is similar: however, the *load_training* parameter is absent, as the PDX data can be used only as testing.

```{r}
sldata <- load_prepared_pdx_data(confident = 'conf',
                                 uniformed = TRUE,
                                 fs_type = c('bio_driven','lasso'),
                                 type = 'sl',
                                 samples_filter = NULL)
```

# Values for *fs_type*

A table with values for *fs_type* is reported, together with its meaning:

| Value      | Description |
| ----------- | ----------- |
| c('ntp_only')      | CRIS signature (564 genes) |
| c('bio_driven')   | Genes surviving biological driven filter (7100 genes).   |
| c('bio_driven','lasso')      |  Genes surviving biological driven filter and union of 10 lasso filters (236 genes). |
| c('bio_driven','ntp_only')   | CRIS signature genes surviving biological driven filter (178 genes). |
| c('bio_driven','lasso','ntp_only')   | CRIS signature genes surviving biological driven filter and lasso (12 genes).|