---
title: "How to filter data"
output: html_notebook
---

# Setup

In order to load the data, the *here* library is used to access the path with respect to the working directory of the project (CRIS_single-sample folder, which contains the .Rproj file). In order to apply the filters, we first need to load some data.

```{r setup}
library(here)
source(here('src','load_libraries.r'))
source(here('src','utils','source_utils.r'))
source(here('src','loader_writer','load_data.r'))
source(here('src','data_management','source_data_management.r'))

# Data loading
load_candiolo_hg19()
```

# Filter by genes

Genes can be filtered by list (extract only some of the genes) or by expression threshold(for example, extract the genes for which the expression is at least 6 for at least 30\% of the samples). In both cases, we need to instantiate a gene filter:

```{r create_gene_filter}
gf <- GenesFilter$new()
genes_selection <- c("A1BG",  "A1CF",  "A2BP1")
```

The expression matrix can be extracted from the data ExpressionSet as follows:

```{r extract_exp_matr}
exp_matr <- CANDIOLO_HG19@assayData$exprs
exp_matr[5:7, 5:7]
```

The expression matrix contains genes on the rows and samples on the columns. In order to filter genes by a predefined list, it is sufficient to:

```{r gene_filt_by_list}
data_filt_by_list <- gf$filter_by_list(exp_matr, genes_selection)
data_filt_by_list
```

A message is displayed with the genes selected by the filter.

In order to filter the genes by expression, it is sufficient to:

```{r filt_by_threshold}
data_filt_by_exp <- gf$filter_by_threshold(exp_matr, expr_thr = 6, samples_thr = 0.3)
```

As before, a message is displayed with the selected number of genes. In both cases, an expression matrix is returned. To build an ExpressionSet with the new matrix it is sufficient to:

```{r build_expset}
filt_expset <- ExpressionSet(assayData = data_filt_by_exp)
```

This is quite relevant as data are mostly saved and used as ExpressionSets.

# Filter by samples

The same expression matrix can be filtered also by samples. As before, the filter must be instantiated.

```{r create_samples_filter}
sf <- SamplesFilter$new()
samples_selection <- c("TCGA-AF-2693-01A-02R-1736-07", "TCGA-AA-3495-01A-01R-1410-07")
```

In order to filter the samples, it is sufficient to:

```{r apply_samples_filter}
sf$filter_by_list(exp_matr, samples_selection)
```

As for the genes filter, the filter displays an information on the number of samples extracted.
It is also possible to take all samples but the ones specified:

```{r remove_samples_filter}
sf$remove_by_list(exp_matr, samples_selection)
```

# Combining filters

The filters can also be combined:

```{r combine_filters}
filt_data_combined <- sf$filter_by_list(exp_matr, samples_selection)
filt_data_combined <- gf$filter_by_list(filt_data_combined, genes_selection)
```












