---
title: "Reference results"
output: html_notebook
---

# Setup

In order to load the data, the *here* library is used to access the path with respect to the working directory of the project (CRIS_single-sample folder, which contains the .Rproj file). In order to apply the filters, we first need to load some data.

```{r setup}
library(here)
source(here('src','load_libraries.r'))
source(here('src','utils','source_utils.r'))
source(here('src','loader_writer','load_data.r'))
source(here('src','data_management','source_data_management.r'))

# Data loading
load_gmql_grch38(filtered = TRUE, uniformed = TRUE)
```

# Published results

Loading of published results for NTP and TSP (TCGA\_HG19 dataset)

```{r}
load_published_ntp()
load_published_tsp()
```

# Training and testing samples

TCGA\_GRCH38 dataset has been splitted into training and testing:

```{r}
training_samples <- load_training_samples()
```

```{r}
testing_samples  <- load_tcga_testing_samples()
```

# NTP reference result

The reference NTP results contain the data used for the NTP classification (*data*), the result of the classification (*result*) and the number of samples assigned to each class (*summary*).

```{r}
NTP_ref <- load_file(path_loader$get_path("NTP_REF_TCGA"))
PDX_ref <- load_file(path_loader$get_path("NTP_REF_PDX"))
```





