---
title: "NTP replication"
output: html_notebook
---

# Setup

In order to load the data, the *here* library is used to access the path with respect to the working directory of the project (CRIS_single-sample folder, which contains the .Rproj file). 

```{r setup}
library(here)
source(here('src','load_libraries.r'))
source(here('src','utils','source_utils.r'))
source(here('src','loader_writer','load_data.r'))
source(here('src','data_management','source_data_management.r'))
source(here('src','classifiers','source_classifiers.r'))
source(here('src','pipelines','source_pipelines.r'))
source(here('src','biological_validation','source_biological_validation.r'))

```

**NB**: For the replication of NTP classifier, actually, *n = 100,000 resamplings* were used. Here, we put a smaller number of resamplings (*100*) to allow a faster application of the classifier. It is appropriate to use at least *1000* resamplings to obtain a reliable value for BH.FDR scores.

# Replication on hg19 data (with published reference)

In order to apply the replication on original hg19 data, it is sufficient to:

```{r hg19_replication}
# Get the data
load_candiolo_hg19()
crisdata <- cris_no_ref_data_pipeline(expset = CANDIOLO_HG19)

# Get the NTP features
load_features_hg19(original = TRUE)

# Get the NTP reference
ntp_ref_hg19 <- load_published_ntp()

# Replicate the NTP classification
ntp_res_hg19 <- ntp_pipeline(crisdata = crisdata,
                             features = features_hg19_original,
                             repl_id = 'replication_hg19_original',  # name to distinguish different replications
                             ntp_ref = ntp_ref_hg19,
                             n_resempl = 100)

```

The result contains the obtained prediction and a comparison with the provided reference. 

# NTP replication (no reference)

The reference NTP classification has been obtained with the following parameters:

```{r tcga_grch38_replication}
# Get the data
ntp_data <- load_prepared_tcga_data(confident = 'all',       # Take all the samples
                                    uniformed = TRUE,
                                    fs_type = 'none',        # Do not filter genes
                                    type = 'cris_no_ref')
# Get the NTP features
load_features_grch38(dataset = 'tcga')

# Replicate the NTP classification
ntp_res <- ntp_pipeline(crisdata = ntp_data,
                        features = features_grch38,
                        repl_id = 'replication_all',         # name to distinguish different replications
                        n_resempl = 100000)

```

Again, we used 100000 resamplings, but here we indicate 100 only to allow a faster computation. The same procedure can be applied to classify merged PDX data too.

# Reference classification

The reference classifications can be loaded as:

```{r}
ntp_ref_tcga <- load_file(path_loader$get_path('NTP_REF_TCGA'))
ntp_ref_pdx  <- load_file(path_loader$get_path('NTP_REF_PDX'))
```

The biological validation will be conducted on data used for testing, both in TCGA and in PDX. Because of this, we need to load the proper data objects as follows:

```{r data_for_ntp_bio_validation}
sldata <- load_prepared_tcga_data(confident = 'conf',
                                   uniformed = TRUE,
                                   fs_type = 'ntp_only',
                                   type = 'sl')

mldata <- load_prepared_tcga_data(confident = 'conf',
                                 uniformed = TRUE,
                                 fs_type = 'ntp_only',
                                 type = 'ml')

```

We will use the *test_ref* field of each data object to obtain kaplan meier curves and fisher tests.

# Kaplan Meier curves

In order to plot the Kaplan-meier curve, we need first to prepare the data and then computing the curve. Finally, we can use the obtained result to show the plot.

```{r kaplan_meier_ntp_sl}
# Object that handles the Kaplan Meier computation
annot <- load_file(path_loader$get_path('GMQL_GRCH38_ANNOT'))
biopl <- BiologicalPlots$new(annotations = annot, xlim = 36)

# Prepare the data for Kaplan Meier
km_ntp_sl_ref <- biopl$ref_kaplan_meier(sldata$test_ref, type = 'ntp_sl', cl = 'CRIS-B')

# Compute the Kaplan Meier
km_ntp_sl <- biopl$compute_kaplan_meier(km_ntp_sl_ref, group = 'ref',cl = 'CRIS-B')

# Show the plot
biopl$show_km_plot(km_ntp_sl, alg_name = 'ntp sl')
```

Similarly, for multi-label NTP:

```{r kaplan_meier_ntp_ml}
# Object that handles the Kaplan Meier computation
annot <- load_file(path_loader$get_path('GMQL_GRCH38_ANNOT'))
biopl <- BiologicalPlots$new(annotations = annot, xlim = 36)

# Prepare the data for Kaplan Meier
km_ntp_ml_ref <- biopl$ref_kaplan_meier(mldata$test_ref, type = 'ntp_ml', cl = 'CRIS-B')

# Compute the Kaplan Meier
km_ntp_ml <- biopl$compute_kaplan_meier(km_ntp_ml_ref, group = 'ref',cl = 'CRIS-B')

# Show the plot
biopl$show_km_plot(km_ntp_ml, alg_name = 'ntp ml')
```

# Biological validation

In order to perform the biological Fisher tests, it is sufficient to create an annotation object (TCGA or PDX, depending on the dataset) and, after having read the biological annotations, simply call two functions:

```{r tcga_bio_validation_ntp_sl}
# Create the annotation object and read the annotations
ann      <- AnnotationTCGA$new()
annot    <- load_file(path_loader$get_path('GMQL_GRCH38_ANNOT'))

# Get count of each variable (used to build contingency table of fisher tests) and perform fisher tests
count_ntp_sl <- ann$get_count_annot_sl(sldata$test_ref, annot, 'ntp_sl')
tests_ntp_sl <- ann$perform_stat_tests(count_ntp_sl)
```

In order to perform the same tests on ntp multi-label:

```{r tcga_bio_validation_ntp_ml}
# Get count of each variable (used to build contingency table of fisher tests) and perform fisher tests
count_ntp_ml <- ann$get_count_annot_ml(res = NULL, annot, mldata$ref, type = 'ntp_ml')
tests_ntp_ml <- ann$perform_stat_tests(count_ntp_ml) 
```

In order to save the data to draw forest plots in excel:

```{r forest_plots}
biopl <- BiologicalPlots$new(annotations = annot)
ntp_sl_forest_plots <- biopl$compute_all_forest_plots(.FOREST_ATTRIBUTES$tcga, 
                                                      .FOREST_LABEL_ATTRIBUTES$tcga, 
                                                      .FOREST_CLASSES, 
                                                      tests_ntp_sl)

ntp_ml_forest_plots <- biopl$compute_all_forest_plots(.FOREST_ATTRIBUTES$tcga, 
                                                      .FOREST_LABEL_ATTRIBUTES$tcga, 
                                                      .FOREST_CLASSES, 
                                                      tests_ntp_ml)
```

The data so obtained can be copied into the forest plot template (present in the additional files folder) following the structure already present in the template. The plots will update automatically. The only thing to be updated manually in the plots is the text of the etiquette of each test (as they represent p-values). Finally, the tests for multi-label classifier should be divided into primary class and all samples assigned to the class.
