---
title: "Single-label replication"
output: html_notebook
---

# Setup

In order to load the data, the *here* library is used to access the path with respect to the working directory of the project (CRIS_single-sample folder, which contains the .Rproj file). 

```{r setup}
library(here)
source(here('src','load_libraries.r'))
source(here('src','utils','source_utils.r'))
source(here('src','loader_writer','load_data.r'))
source(here('src','data_management','source_data_management.r'))
source(here('src','classifiers','source_classifiers.r'))
source(here('src','pipelines','source_pipelines.r'))
source(here('src','biological_validation','source_biological_validation.r'))
```

# Data loading

The single-label classifiers are trained on training TCGA data samples. The data can be loaded with:

```{r sldata}
sldata <- load_prepared_tcga_data(confident = 'conf',
                                  uniformed = TRUE,
                                  fs_type = 'ntp_only',
                                  type = 'sl',
                                  load_training = TRUE)
```

The structure of training data (accessed with *train_* field) and its corresponding reference (*train_ref* field) can be inspected with:

```{r train_data_structure}
head(sldata$train_)
head(sldata$train_ref)
```

# The generic single-label classifier model

The single-label classifier models are actually functions that describe the class label as function of the feature genes. Therefore, the formula to model the classifiers is:

```{r sl_formula}
.SL_FORMULA
```

Obviously, the class label is included in the training data.

```{r check_train_data}
CLASS_LABEL %in% colnames(sldata$train_)
```

The methods that can be used to train single-label classifiers are hold in:

```{r sl_available_models}
methods
```

# Training with default hyperparameter tuning

The formula described above is used within the library *caret* to model the single-label classifier. In order to train the model with default tuning (provided by caret).

```{r default_training}
sl_train <- sl_pipeline_train(sldata = sldata,
                              method = methods[1],           # linear SVM, one of the applied models
                              sl_formula = .SL_FORMULA,
                              seed = .SEED)
```

The obtained model can be accessed with:

```{r lsvm_default_tuning}
lsvm_default <- sl_train$model
```

# Training with custom tuning

It is possible to tune the model with custom tuning values, which are defined in *src/classifiers/classifiers_settings.r*. In order to do this, it is sufficient to provide a *trainControl* settings (defined in the above script as *fit_control*) and a list of tuning values, which can be accessed using the method name as follows:

```{r lsvm_tuning_grid}
# Linear SVM tuning values
tune_grid[[methods[1]]]
```

The function to be called in order to train a linear SVM with custom tuning is the following:

```{r custom_tuning_function}
sl_train_tuned <- sl_pipeline_train(sldata = sldata,
                                    method = methods[1],      # linear SVM, one of the applied models
                                    sl_formula = .SL_FORMULA,
                                    seed = .SEED,
                                    tune = TRUE,
                                    fit_control = fit_control,
                                    tune_grid = tune_grid[[methods[1]]])
```

As before, the model can be extracted with:

```{r custom_tuned_lsvm}
lsvm_custom <- sl_train_tuned$model
```

# Testing of the model

In order to test the obtained model, it is possible to use a custom testing pipeline, providing both the sldata and the obtained model.

```{r testing_default_model}

lsvm_default_testing <- sl_pipeline_test(sldata = sldata,
                                         method = methods[1],
                                         seed = .SEED,
                                         model = lsvm_default)

```

The performance of the classification can be visualized with

```{r performance_default_model}

print(lsvm_default_testing$metrics)

```

Similarly, for the custom model:

```{r testing_custom_model}

lsvm_custom_testing <- sl_pipeline_test(sldata = sldata,
                                         method = methods[1],
                                         seed = .SEED,
                                         model = lsvm_custom)
print(lsvm_custom_testing$metrics)
```

# Kaplan Meier curves

In order to plot the Kaplan-meier curve, we need first to prepare the data and then computing the curve. Finally, we can use the obtained result to show the plot.

```{r}
# Object that handles the Kaplan Meier computation
annot <- load_file(path_loader$get_path('GMQL_GRCH38_ANNOT'))
biopl <- BiologicalPlots$new(annotations = annot, xlim = 36)

# Prepare the data for Kaplan Meier
ref   <- lsvm_custom_testing$binary_res %>% rownames_to_column(ALIQUOT_LABEL)
km_lsvm_ref <- biopl$ref_kaplan_meier(ref, type = 'sl', cl = 'CRIS-B')

# Compute the Kaplan Meier
km_lsvm <- biopl$compute_kaplan_meier(km_lsvm_ref, group = 'ref',cl = 'CRIS-B')

# Show the plot
biopl$show_km_plot(km_lsvm, alg_name = 'lsvm (single-label)')
```


# Biological validation

In order to perform the biological Fisher tests, it is sufficient to create an annotation object (TCGA or PDX, depending on the dataset) and, after having read the biological annotations, simply call two functions:

```{r tcga_bio_validation}
# Create the annotation object and read the annotations
ann      <- AnnotationTCGA$new()
annot    <- load_file(path_loader$get_path('GMQL_GRCH38_ANNOT'))

# Get count of each variable (used to build contingency tabel of fisher tests) and perform fisher tests
count_sl <- ann$get_count_annot_sl(lsvm_custom_testing, annot, 'sl')
tests_sl <- ann$perform_stat_tests(count_sl)
```

The *perform_stat_tests* methods prints the result of each test, providing in green tests with a significative p-value (< 0.05).

Obviously, the tests can be performed also on PDX (in this case, only test for **Cetuximab sensitivity** is available). In order to execute them, simply test the classifiers on PDX data, create an *AnnotationPDX* object and, after having read the correct annotation data, call the same functions:

```{r pdx_testing_and_bio_validation}
# Read PDX data
sldata_pdx <- load_prepared_pdx_data(confident = 'conf',
                                     uniformed = TRUE,
                                     fs_type = 'ntp_only',
                                     type = 'sl')
# Apply the classifier on PDX data
lsvm_custom_testing_pdx <- sl_pipeline_test(sldata = sldata_pdx,
                                            method = methods[1],
                                            seed = .SEED,
                                            model = lsvm_custom)
print(lsvm_custom_testing_pdx$metrics)

# Create the annotation object and read the annotations
ann_pdx   <- AnnotationPDX$new()
annot_pdx <- load_file(path_loader$get_path('PDX_MERGED_ANNOT'))

# Get count of each variable (used to build contingency table of fisher tests) and perform fisher tests
count_sl_pdx <- ann_pdx$get_count_annot_sl(lsvm_custom_testing_pdx, annot_pdx, 'sl')
tests_sl_pdx <- ann_pdx$perform_stat_tests(count_sl_pdx) 
```

In order to save the data to draw forest plots in excel:

```{r forest_plots}
biopl <- BiologicalPlots$new(annotations = annot)
sl_forest_plots <- biopl$compute_all_forest_plots(.FOREST_ATTRIBUTES$tcga, 
                                                  .FOREST_LABEL_ATTRIBUTES$tcga, 
                                                  .FOREST_CLASSES, 
                                                  tests_sl)
```

The data so obtained can be copied into the forest plot template (present in the additional files folder) following the structure already present in the template. The plots will update automatically. The only thing to be updated manually in the plots is the text of the etiquette of each test (as they represent p-values).


