---
title: "TSP replication"
output: html_notebook
---

# Setup

In order to load the data, the *here* library is used to access the path with respect to the working directory of the project (CRIS_single-sample folder, which contains the .Rproj file). 

```{r setup}
library(here)
source(here('src','load_libraries.r'))
source(here('src','utils','source_utils.r'))
source(here('src','loader_writer','load_data.r'))
source(here('src','data_management','source_data_management.r'))
source(here('src','classifiers','source_classifiers.r'))
source(here('src','pipelines','source_pipelines.r'))
source(here('src','biological_validation','source_biological_validation.r'))

```

# TSP replication 
In order to apply the TSP classifier, it is sufficient to apply the pipeline on *CRISData* object.

```{r confident_testing_cris_signature}

tsp_data <- load_prepared_tcga_data(confident = 'conf',
                                    uniformed = TRUE,
                                    fs_type = 'ntp_only',
                                    type = 'cris',
                                    samples_filter = load_tcga_testing_samples()
                                    )
tsp_res <- tsp_pipeline(tsp_data)
print(tsp_res$metrics)
```

In this case, the pipeline has been applied *only on the 169 TCGA data used for testing in single-label and multi-label alternative classifiers*. In order to apply the TSP classifier on the entire TCGA dataset:

```{r all_samples_cris_signature}
tsp_all <- load_prepared_tcga_data(confident = 'all',
                                    uniformed = TRUE,
                                    fs_type = 'ntp_only',
                                    type = 'cris')

tsp_res_all <- tsp_pipeline(tsp_all)
print(tsp_res_all$metrics)
```

NB: the only accepted value for *fs_type* is ntp_only. Other values will give error during the execution of the pipeline:

```{r other_signature, eval=FALSE, include=FALSE}
tsp_bio <- load_prepared_tcga_data(confident = 'all',
                                    uniformed = TRUE,
                                    fs_type = c('bio_driven','ntp_only'),
                                    type = 'cris')
tsp_pipeline(tsp_bio)

```


# Kaplan Meier curves

In order to plot the Kaplan-meier curve, we need first to prepare the data and then computing the curve. Finally, we can use the obtained result to show the plot.

```{r}
# Object that handles the Kaplan Meier computation
annot <- read_excel(path_loader$get_path('GMQL_GRCH38_ANNOT'))
biopl <- BiologicalPlots$new(annotations = annot, xlim = 36)

# Prepare the data for Kaplan Meier
ref   <- tsp_res$binary_res %>% rownames_to_column(ALIQUOT_LABEL)
km_tsp_ref <- biopl$ref_kaplan_meier(ref, type = 'sl', cl = 'CRIS-B')

# Compute the Kaplan Meier
km_tsp <- biopl$compute_kaplan_meier(km_tsp_ref, group = 'ref',cl = 'CRIS-B')

# Show the plot
biopl$show_km_plot(km_tsp, alg_name = 'tsp')
```


# Biological validation

In order to perform the biological Fisher tests, it is sufficient to create an annotation object (TCGA or PDX, depending on the dataset) and, after having read the biological annotations, simply call two functions:

```{r tcga_bio_validation}
# Create the annotation object and read the annotations
ann      <- AnnotationTCGA$new()
annot    <- read_excel(path_loader$get_path('GMQL_GRCH38_ANNOT'))

# Get count of each variable (used to build contingency tabel of fisher tests) and perform fisher tests
count_tsp <- ann$get_count_annot_sl(tsp_res, annot, 'sl')
tests_tsp <- ann$perform_stat_tests(count_tsp)
```


```{r pdx_testing_and_bio_validation}
# Read PDX data
crisdata_pdx <- load_prepared_pdx_data(confident = 'conf',
                                     uniformed = TRUE,
                                     fs_type = 'ntp_only',
                                     type = 'cris')
# Apply the classifier on PDX data
tsp_pdx <- tsp_pipeline(crisdata_pdx)
print(tsp_pdx$metrics)

# Create the annotation object and read the annotations
ann_pdx   <- AnnotationPDX$new()
annot_pdx <- read_excel(path_loader$get_path('PDX_MERGED_ANNOT'))

# Get count of each variable (used to build contingency table of fisher tests) and perform fisher tests
count_tsp_pdx <- ann_pdx$get_count_annot_sl(tsp_pdx, annot_pdx, 'sl')
tests_tsp_pdx <- ann_pdx$perform_stat_tests(count_tsp_pdx) 
```

In order to save the data to draw forest plots in excel:

```{r forest_plots}
biopl <- BiologicalPlots$new(annotations = annot)
tsp_forest_plots <- biopl$compute_all_forest_plots(.FOREST_ATTRIBUTES$tcga, 
                                                  .FOREST_LABEL_ATTRIBUTES$tcga, 
                                                  .FOREST_CLASSES, 
                                                  tests_tsp)
```
